Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-16 paper/structureScene.tex
clarify between bio and eco acoustics

An important task in computational ecoacoustics is to determine the similarity of two audio scenes. This is often done using bag-of-frames models and other representations based on summary statistics. While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background. To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure. To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs). Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.

@@

% Note that keywords are not normally used for peerreview paper.
\begin{IEEEkeywords}
ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
\end{IEEEkeywords}

@@

This works falls in the realm of ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
Consequently, few well-designed datasets are readily available for evaluation purposes.

@@

This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales. Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification. In order to extend the ideas of late integration to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in ecoacoustics.

>>>>

An important task in the emerging field of computational ecoacoustics is to determine the similarity of two audio scenes to be able to get a high level understanding of large amount of audio recordings coming from large deployment of acoustics sensors that are continuously recordings the sonic environment. A convenient method considers bag-of-frames models and other representations based on summary statistics. While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background. To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure. To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs). Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.

@@

% Note that keywords are not normally used for peerreview paper.
\begin{IEEEkeywords}
acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
\end{IEEEkeywords}

@@

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
Consequently, few well-designed datasets are readily available for evaluation purposes.

@@

This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales. Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification. In order to extend the ideas of late integration to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in computational bioacoustics.
